---
title: "Lesonia trabeculata data"
author: "SUBELAB"
date: "26 de agosto de 2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(kableExtra)
```


1. IFOP Database

1.1. Sites & georeferences by year
```{r}
coord<-read.csv("coordenadas.csv")
 kable(coord) %>% kable_styling(bootstrap_options = c("striped", 
    "hover", "condensed"), full_width = F)
unique(coord$Site)
```
1.2. Databases
This project evaluate the kelp forest state in TURF sites (also called AMERB in Spanish) evaluating kelp density, kelp morphology, herbivores and fish communities. 

  - Ifop_Kelp_Density= To evaluate kelp density, 7-10 transects of 100x2 meters were monitored in each site. The transect is divided into 10 m stations (n=10), where adult (>10 cm holdfast) and juvenile (<10 cm holdfast) kelp were counted at 1 m on each side (20 m^2).

```{r}
IFOP_Kelp_Density<- read_csv("IFOP_Kelp_density.csv") %>% select (c(1:8,13:18,21))%>% mutate(Site=as.factor(Site),Management=as.factor(Management),Species=as.factor(Species), Stage=as.factor(Stage)) %>% filter(Species=="Lessonia trabeculata")
summary(IFOP_Kelp_Density)

```


  - Ifop_Kelp_Morpho= In 3 of the transects, the morphology (holdfast size, no. stipes, total lenght and frondosity) and herbivores density are quantified in 1m^2 quadrants.Between 7-10 quadrants were spread equidistant across the transect.


```{r}
#Only kelp information
IFOP_Kelp_Morpho<- read_csv("IFOP_Kelp_Morphology.csv") %>% select(c(1:9,11:13,15:19)) %>%  mutate(Site=as.factor(Site),Management=as.factor(Management),Group=as.factor(Group),Species=as.factor(Species)) %>% filter(Group=="Macroalga", Species =="Lessonia trabeculata")
summary(IFOP_Kelp_Morpho)


```
  
  - Ifop_allometry= In 2018, ~30 algae were extracted from 6 of 8 monitored sites, and total weight, holdfast diameter, holdfast perimeter, number of stipes, and total lenght were measured.
  
```{r}
IFOP_allometry = read.csv("IFOP_allometry.csv") %>% select(-16)
summary(IFOP_allometry)
```
  


2. NERC Database

A harvesting experiment was carried out in the North, Center-North and Center of Chilean coast, comparing Open Access Areas (OA) and Management Areas (TURF). An initial monitoring was carried out, prior to the extraction (Baseline- bl) and a monitoring after 12, 18, 24 and 32 months (MAE in databases).

2.1 Baseline : Same protocol than IFOP
    - NERC_Kelp_Density_bl= 7-10 transects of 100x2 meters were divided into 10 m stations (n=10), where adult (>10 cm holdfast) and juvenile (<10 cm holdfast) kelp were counted at 1 m on each side (20 m^2).
    - NERC_Kelp_Morpho_bl= In 3 of the transects, the morphology and herbivores density were quantified in 1m^2 quadrants.Between 7-10 quadrants were spread equidistant across the transect.

```{r}
NERC_Kelp_Density_bl= read.csv("Kelp_transect.csv")
summary(NERC_kelpdensity_bl)

NERC_Kelp_Morpho_bl= read.csv("Quads_baseline.csv") %>% filter(Species=="Lessonia trabeculata")
summary(NERC_Kelp_Morpho_bl)
```


2.2. Post-harvesting Monitoring : 
  - NERC_Kelp_Density_ph
  - NERC_Kelp_Morpho_ph
  
```{r}
NERC_Kelp_Density_ph=read.csv("Patches_quad.csv")
summary(NERC_Kelp_Density_ph)
NERC_Kelp_Morpho_ph=read.csv("Patches_Morphology.csv")
summary(NERC_Kelp_Morpho_ph)
```
  

3. Nerc_Cata's thesis

3.1 Allometry
  - 12-19 kelp were harvested in NERC sites, covering all the size range.
  - Weight and morphology were measured, and the relation FW:DW and Carbon content was also estimated. 
  - Measurements in Stipes (First_bif, Diam_stipes, LT_stipes, Dichot_stipes) are a mean between 3  stipes
  
  The variables measured in each individuals were:
    1. Total.Weight: Total weight of algae (cm)
    2. LT: Total lenght of algae (cm)
    3. Diam_Holdfast: Maximum diameter of holdfast (cm)
    4. High_Holdfast: Holdfast height (cm)
    5. Weight_Holdfast: Holdfast weigth (g)
    6. Vol_Holdfast: Holdfast volume (ml)
    7. Estipes: Stipe numbers 
    8. First_bif: Length from the beginning of stipe to the first dichotomy (cm)
    9. Diam_stipes: In the middle of the stipe, under the first dichotomy (mm)
    10. LT_stipes: From the start of the stipe on the holdfast to the last dichotomy (cm)
    11. Dichot_stipes: Number of dichotomies from the start of the stipe to the last dichotomy
    12. Weight_stipes: Weight of all stipes (g)
    13. Reproductive: Presence/Absence reproductive blades
    14. Weight blades: Weight of all blades 
  
```{r}
Nerc_Allometry= read.csv("Morfo_DW_C.csv")
summary(Nerc_Allometry)
```

3.2 Growth and survival monitoring
   The growth of kelp after a harvesting event have been monitored in tagged recruits every 6 month in patches of NERC experiment. When kelp disapeared the holdfast value are 0.
   
```{r}
Growth<- read.csv("Growth&Survival.csv")
```

